Supplemental Information: Aligning marine species range data to better serve science and conservation

### 2016-04-02 manuscript-main.Rmd
submit this bastard

_Keywords:_ <!--Keywords are listed below the abstract of the published article. At least three keywords are required at submission.  maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> TBD: These should be words NOT in the title or abstract (BSH comment)

>>>>

_Keywords:_ <!--Keywords are listed below the abstract of the published article. At least three keywords are required at submission.  maybe species distribution maps, AquaMaps, IUCN, marine protected area planning, biodiversity status--> Marine spatial planning, species distribution model, extent of occurrence, spatial ecology

### 2016-04-02 ms_2_intro_res_disc.Rmd
submit this bastard

## How and why the datasets differ

The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (2) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.  In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.  The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global raster of "probability of occurrence" for each species (14, 15).  Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (3, 4).

The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).  Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.  Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (12).  By comparing maps resulting from IUCN and AquaMaps methodolog, we can identify and possibly address mechanistic causes for each type of error.

@@

The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area. These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig. 2B). This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.  We expected to see significant concurrence between maps for IUCN Red List species of concern, assuming that increased attention would lead to improved understanding of species ranges; however, we found no compelling patterns to indicate that this was the case (Fig. S2).

@@

Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.  Consequently, area ratios are close to 100%, though similar areas are meaningless when the distributions are poorly aligned.

@@

The global Ocean Health Index (OHI) (3, 4), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones. To maximize the number of represented species, OHI gleans spatial distributions from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.  OHI currently uses a probability of occurrence threshold of 40% to determine species presence for AquaMaps data.

@@

Klein et al. (11) compare the global distribution of species to the global distribution of marine protected areas to assess how well the MPAs protect species and identify which species fall through gaps in protection.  The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.  They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all (i.e., "gap" species).

We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using 2015 AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig. 4</a>).  We found a five-fold increase in the proportion of gap species (6.4% of species vs. 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs. 47.7%).  However, this comparison also indicates a larger proportion of well-protected species with greater than 10% of range protected (2.9% of species vs. 1.5%).  Conclusions drawn from each of these datasets would paint dramatically different pictures of the protection afforded by our current global MPA network.

>>>>

The IUCN publishes species range maps based on expert input of spatial boundaries of a given species' "limits of distribution" (2) - essentially a refined extent of occurrence, based on observation records and informed by expert understanding of species' range and habitat preferences.  In contrast, AquaMaps models species distribution based on environmental preferences (e.g., temperature, depth, salinity) deduced from occurrence records, published species databases such as FishBase, and expert knowledge.  The AquaMaps model overlays these environmental preferences atop a map of environmental attributes on a 0.5 degree grid, creating a global map of "probability of occurrence" for each species (14, 15).  Studies using AquaMaps data frequently define "presence" by applying a probability threshold, e.g., the Ocean Health Index defines presence as 40% or greater probability of occurrence (3, 4).

The methodologies behind the creation of these datasets imply differences in prediction of species distribution due to errors of commission (falsely indicating species presence) and omission (falsely indicating species absence).  Geographic range data such as IUCN range maps frequently introduce commission errors, while species distribution models such as AquaMaps will likely introduce fewer commission errors but more omission errors.  Each type of error bears different implications for conservation goals: commission errors can result in prioritizing areas not relevant to conservation goals, while omission errors may result in protected area networks that fail to include important habitat and range (12).  By comparing maps resulting from IUCN and AquaMaps methodologies, we can identify and possibly address mechanistic causes for each type of error.

@@

The upper right quadrant includes species (n = `r n_q1`) whose described ranges are above average in alignment of both spatial distribution and area. These species tend to be well-studied and include wide-ranging pelagic organisms such as marine mammals, tunas, and billfishes (Fig. 2B). This result is not surprising, as species with very large ranges are likely to be more aligned regardless of methodology simply because their ranges span nearly the entire map.

@@

Species found in the lower right quadrant (n = `r n_q3`) often represent cases of "two wrongs make a right."  For these species, IUCN ranges frequently overextend into unsuitable depths, as in the case of many upper left quadrant species, while at the same time AquaMaps ranges often aggressively extrapolate presence into locations where IUCN predicts absence, introducing additional commission errors.  Consequently, area ratios are close to 100%, though similar areas are unhelpful when the distributions are poorly aligned.

@@

The global Ocean Health Index (OHI) (3, 4), a composite index comprising ten sustainable benefits provided by a healthy ocean, uses species spatial distribution data and IUCN Red List conservation status to calculate biodiversity status (scored from zero to 100) for each of the world's 221 exclusive economic zones. To maximize the number of represented species, OHI gleans spatial distribution data from Red List-assessed species in both IUCN and AquaMaps datasets (n = `r format(n_spp_ohi, big.mark = ',')`), prioritizing IUCN data for the `r format(n_spp_both_ohi, big.mark = ',')` species included in both sources.  OHI currently uses a probability of occurrence threshold of 40% to determine species presence for AquaMaps data.

@@

Klein et al. (11) compare the global distribution of species to the global distribution of marine protected areas to assess how well current MPAs overlap with species ranges and identify which species fall through gaps in protection.  The study relied on the AquaMaps database, using a probability of occurrence threshold of 50% or greater, to determine species presence, and the World Database of Protected Areas to define zones of marine protection.  They found that the global MPA network leaves 90.5% of marine species with less than 5% of their overall range represented within MPAs, and 1.4% of species have no protection at all (i.e., "gap" species).

We recalculated the amount of under-protected and gap species using either IUCN or AquaMaps data (using the most recent AquaMaps data and a 0% threshold to allow the most meaningful comparison to IUCN's "limits of distribution", <a href = '#fig4'>Fig. 4</a>).  We found a five-fold increase in the proportion of gap species (6.4% of species vs. 1.2%) and dramatically larger proportion of species with less than 2% of their range protected (73.2% of species vs. 47.7%).  However, this comparison also indicates a larger proportion of well-protected species with greater than 10% of range protected (2.9% of species vs. 1.5%).  Conclusions drawn from each of these datasets would paint dramatically different pictures of the protection afforded by our current global MPA network.

### 2016-04-02 ms_3_methods.Rmd
submit this bastard

__Comparison of paired maps__:
Using genus and species binomials as a matching key, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.  To avoid double-counting, we removed subpopulations and species aliases.  We determined species presence within each spatial cell using the same criteria as outlined above.

@@

For each species with ranges described in both IUCN and AquaMaps, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).  $A_{small \cap large}$ represents the amount of overlapping area between the two datasets. We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.  To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (19) to overlay a selection of IUCN and AquaMaps maps with a 200 meter bathymetry contour.

**Methods for OHI case study: **
Using methods and supplemental materials from OHI (3, 4), we modified the original code for OHI 2015 Species status (SPP), allowing for flexibility in prioritized data source and AquaMaps presence threshold.  We ran the SPP code three times, prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.  We compared each of these to the output of the published OHI 2015 SPP model (which prioritizes IUCN over AquaMaps at a 40% threshold).

>>>>

__Comparison of paired maps__:
Using genus and species binomials as a matching key, we selected the subset of marine species that have range maps in both IUCN and AquaMaps current native distribution.  To avoid double-counting, we removed subpopulations and species aliases.  We determined species presence within each spatial cell using the same criteria outlined above.

@@

For each species with ranges described in both IUCN and AquaMaps, $A_{small}$ and $A_{large}$ indicate the smaller and larger range representation (regardless of which dataset).  $A_{small \cap large}$ represents the amount of overlapping area between the two datasets. We visually inspected a random selection of paired distribution maps from each quadrant to identify possible mechanistic causes of misalignment.  To verify that IUCN predicted unsuitable habitat for depth limited species, we used QGIS (19) to overlay a selection of IUCN and AquaMaps maps with a 200-meter bathymetry contour.

**Methods for OHI case study:**
Using methods and supplemental materials from OHI (3, 4), we modified the original code for OHI 2015 Species status (SPP), allowing for flexibility in prioritized data source and AquaMaps presence threshold.  We ran the SPP code three times, prioritizing IUCN over AquaMaps for a 0% threshold, and prioritizing AquaMaps over IUCN for both a 40% and 0% threshold.  We compared each of these to the output of the published OHI 2015 SPP model (which prioritizes IUCN over AquaMaps at a 40% threshold).

### 2016-04-02 ms_supplemental.Rmd
submit this bastard

---
title: "Supplemental Information"
output:
  word_document: default
  pdf_document:
    number_sections: no
    toc: no
  html_document:
    highlight: haddock
    number_sections: no
    theme: spacelab
    toc: no
---

@@

Text: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).
Figures: Provide a brief legend for each supporting figure after the supporting text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text. When saving TIFF files, use only LZW compression; do not use JPEG compression. Do not save figure numbers, legends, or author names as part of the image. Composite figures must be preassembled.
3D figures: Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
Tables: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.
Datasets: Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
Movies: Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.
Still images: Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
Appendices: PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.
-->

<!--## Data processing - should this go in methods??? -->

<!--
All code is publicly available and freely downloadable [here on GitHub](link to github).

@@

@@

Each shapefile contains polygon limits of distribution indicating regions of presence/absence, additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.-->

## S1: sample maps from each quadrant

@@

>  __Fig. S1.__ Representative species maps to illustrate each quadrant from Fig. 2A.  Note that A-D are arranged to match quadrants in Fig. 2A.
(A: well-matched) Kajikia albida
(B: distribution-matched) Conus episcopatus
(C: area-matched) Conus magnificus
(D: poorly-matched) Acanthurus nigroris

## S2: Risk by quadrant

@@

>  __Fig. S2.__ Breaking down the quadrants by IUCN extinction risk categories (Fig. S2), we found little support for our hypothesis that maps for species with higher extinction risk tend to be better aligned between the two datasets, perhaps correlated to increased expert scrutiny.

### S3: Coral depth map

@@

<!--
Oculina varicosa - Ivory tree coral ![oculina_pic.jpg](figures/oculina/oculina_pic.jpg)
-->

>  __Fig. S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour. According to IUCN, _O. varicosa_ can be found to depths of 152 m.

### S4: AquaMaps threshold examination

@@

AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.  Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value.  A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability.

For our comparisons of global distribution of represented biodiversity and spatial alignment between datasets, we considered "present" to be any cell with a non-zero probability of occurrence, to best approximate the "limits of distribution" as indicated by IUCN maps.  To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0% to 100% and calculated the average species range relative to a zero threshold.

@@

>  __Fig. S4.__ AquaMaps distribution map extent remaining after applying a presence threshold.  (A) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of 28% of total range, with a wide distribution in which some species lose nearly all of their apparent range.  (B) Median remaining extent at increments of presence threshold. Dark green ribbon includes 25% to 75% quantiles of remaining species range, while lighter ribbon includes 5% to 95% quantiles.  Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity of apparent species range to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.

@@

<!--
___Red List inclusion:___

![](figures/bar_by_taxa_assessed.png)-->

a few codes to indicate likely problems?

- DC = depth clipping; IUCN extents go beyond reasonable depth for the species
- DX = data excluded; AM excludes some observations that appear to be used for IUCN
- DP = data poor (AquaMaps acknowledges data poor status? or use less than 10 data points as criteria)
- NX = needs evaluation by expert; AM model likely predicts areas where species is unlikely to be found based on observed occurrences - does AM model reflect same as AM suitable habitat?

Q2:

| species                | quad | group           | codes   | notes |
| -------                | ---- | -----           | -----   | ----- |
| Conus episcopatus      | q2   | cones           |  DP, DX | d-match around Coral Triangle, but IUCN shows much more around Indian Ocean and S Pacific; also IUCN extents
| Ctenochaetus binotatus | q2   |                 |         | IUCN extents beyond AM
| Naso vlamingii         | q2   | acanthuridae    |         | IUCN extents beyond AM
| Thalassoma purpureum   | q2   | wrasses         |         | IUCN extents beyond AM
| Porites nigrescens     | q2   | corals 2        | DP      | AM limited to CT; IUCN shows far greater dist, and extents
| Conus ammiralis        | q2   | cones           | DP      | AM  | limited to CT, IUCN shows in Madagascar; also extents
| Conus tessulatus       | q2   | cones           |         | extents
| Acropora sarmentosa    | q2   | corals 1        |         | extents; also AM shows HI but IUCN does not?
| Holothuria fuscogilva  | q2   | sea cucumbers   |         | extents; also AM shows Mex and Central Am west coast, but IUCN does not (not available online?)
| Oculina Varicosa       | q2   | corals 2        |         | lim to Caribbean; extents - this might be a good close-up map
| Acanthocybium solandri | q2   | tunas/billfish  |         | mostly tropical by IUCN; AM shows much farther N and S

Q3

| species                | quad | group           | codes   | notes |
| -------                | ---- | -----           | -----   | ----- |
| Conus magnificus       | q3   |  cones          |         | IUCN distribution is more limited than AM distribution)
| Abudefduf concolor     | q3   |  damselfish     |         | Baja for AM, southern Central Am for IUCN; maybe a good closeup map?  AM points from Baja to Peru incl Galapagos; some rejected pts from Caribbean and eastern Atlantic
| Centropyge aurantonotus | q3  |  angelfish      |         | AM shows Caribbean; IUCN shows northern South America, also maybe a good closeup map? AM points from FL to south Brazil, but no Gulf of Mexico.
| Chlorurus perspicillatus | q3 |  parrotfish     |         | IUCN shows only HI; AM shows dots all over S Pacific and near Marianas Islands
| Sarda orientalis        | q3  |  tunas/billfish |         | limited dists by IUCN
| Montastraea franksi     | q3  |  corals 2       | DP      | IUCN shows just in Caribbean; AM shows all up & down east coast of N America and S America

Q4

| species                  | quad | group   | codes   | notes |
| -------                  | ---- | -----   | -----   | ----- |
| Acanthurus nigroris      | q4   |         |         | IUCN dist limited to HI, while AM shows all over S Pacific and CT - AM observations all over S pacific...
| Praealticus tanegasimae  | q4   | blenny  |         | IUCN shows limited dist around Japan/Marianas; AM shows much broader through CT and W Pacific - good local map?
| Stethojulis marquesensis | q4   | wrasse  |         | IUCN shows only small cluster of islands (Marquesas?) while AM shows much broader across S Pacific
| Canthigaster leoparda    | q4   | puffer  |         | IUCN limited spots around a couple of islands; AM shows all throughout CT
| pentacheles snyderi      | q4   | lobster | DP      | AM shows all over - sim to depth profiles? while IUCN shows only a couple patches - HI? S Pacific, and S Indian oceans?  AM limits to specific FAO regions, which results in odd cutoffs; also shows some data points in N Atlantic which has an odd result.
| Nephropsis sulcata       | q4   | lobster |         | IUCN shows odd ranges, most of which don't overlap with AM

-->

<!--
___table to show data poor status and reviewed status for AquaMaps maps represented in the quadrant plot___ _- perhaps update the quadrant plot to reveal data-poor species (and reviewed species? little overlap of reviewed & data-poor) instead of, say, red-list category which doesn't get discussed in the body of the paper?  then this table can go in SOM if we like it_

| quadrant | n species | n data poor species | mean (median) data points | n reviewed species |
| :------: | :-------: | :-----------------: | :-----------------------: | :----------------: |
| all AM   |   22889   |        8749 (38.2%) |          57.1 (16)        |       1296  (5.7%) |
| AM&IUCN  |    2166   |         457 (21.1%) |          89.9 (33)        |        290 (13.4%) |
|   q1     |     401   |          33  (8.2%) |         233.0 (78)        |        100 (24.9%) |
|   q2     |     682   |         151 (22.1%) |          77.4 (39)        |        100 (14.7%) |
|   q3     |     682   |         114 (16.5%) |          52.4 (29)        |         65  (9.5%) |
|   q4     |     410   |         159 (39.7%) |          32.1 (13)        |         25  (6.2%) |

-->

>>>>

---
title: "Supplemental Information: Aligning marine species range data to better serve science and conservation"
author: "Casey O'Hara, Jamie C. Afflerbach, Courtney Scarborough, Benjamin S. Halpern"
output:
  word_document: default
  pdf_document:
    number_sections: no
    toc: no
  html_document:
    highlight: haddock
    number_sections: no
    theme: spacelab
    toc: no
---

@@

Text: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference).
Figures: Provide a brief legend for each supporting figure after the supporting text. Provide figure images in TIFF, EPS, high-resolution PDF, JPEG, or GIF format; figures may not be embedded in manuscript text. When saving TIFF files, use only LZW compression; do not use JPEG compression. Do not save figure numbers, legends, or author names as part of the image. Composite figures must be preassembled.
3D figures: Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
Tables: Supply Word, RTF, or LaTeX files (LaTeX files must be accompanied by a PDF with the same file name for visual reference); include only one table per file. Do not use tabs or spaces to separate columns in Word tables.
Datasets: Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
Movies: Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file. All movies should be submitted at the desired reproduction size and length. Movies should be no more than 10 MB in size.
Still images: Authors must provide a still image from each video file. Supply TIFF, EPS, high-resolution PDF, JPEG, or GIF files.
Appendices: PNAS prefers that authors submit individual source files to ensure readability. If this is not possible, supply a single PDF file that contains all of the SI associated with the paper. This file type will be published in raw format and will not be edited or composed.
-->
<!--
# Data sources and methods

All code is publicly available and freely downloadable [here on GitHub](link to github).

## AquaMaps

@@

## IUCN

@@

Each shapefile contains spatial polygons outlining limits of distribution for each species, indicating regions of presence/absence; additional attributes provide information on extant/extinct ranges, native/introduced ranges, and seasonality.
-->

## Representative maps from each quadrant

@@

>  __Fig. S1.__ Representative species maps to illustrate each quadrant from Fig. 2A.  Note that B-E are arranged to match quadrants in Fig. S1A. (A) Alternative version of Figure 2A that includes extinction risk categories for each species.  Four representative species, one from each quadrant, are mapped in (B-E).
(B) Well-matched: Kajikia albida, the Atlantic white marlin.  Ranges from each data set show nearly complete overlap, and very similar area of coverage.
(C) Distribution-matched: Conus episcopatus, the dignified cone snail.  Ranges shows excellent overlap in the western Pacific, though IUCN range extends beyond the bounds of the AquaMaps range.
(D) Area-matched: Conus magnificus, the magnificent cone snail.  Ranges overlap in the southern Pacific, but align poorly elsewhere.  The total area for each range is similar.
(E) Poorly-matched: Acanthurus nigroris, the blue-lined surgeonfish.  IUCN predicts species range only near the Hawaiian islands; AquaMaps predicts extensive range throughout the central and western Pacific Ocean.  The datasets align in neither distribution nor area of coverage.

## Risk by quadrant

@@

>  __Fig. S2.__ Breakdown of quadrants (Fig. S1A) by IUCN extinction risk categories.  IUCN and Aquamaps ranges were in better agreement (i.e., had a higher probability of being in the upper-right, "well-aligned" quadrant, Figure 2A/Figure S1A) for species at greater risk according to IUCN risk categories (p < 0.001, generalized linear model, family = binomial, Table S1).  This suggests at-risk species may receive more attention in terms of data collection or model refinement.

> __Table S1.__  Generalized linear model (family = binomial) describing how the agreement between IUCN and Aquamap range maps (i.e., probability of occurring in quadrant 1, Figure 2A/S1A) increases for a species as its IUCN risk status increases.  For this analysis, we excluded all species that had "data deficient" IUCN status (n = 255).  We analyzed the IUCN risk categories as a continuous variable (LC = 0.0, NT = 0.2, VU = 0.4, EN = 0.6, CR = 0.8, EX = 1.0).

_Model: glm(quadrant 1 membership ~ IUCN category, family = binomial) n = 2,059_

| Coefficients:  | Estimate | Std. Error | z value | p value      |
| :-----         |   -----: |     -----: |  -----: |       -----: |
| Intercept      | -1.32933 |   0.05804  | -22.902 | < 2e-16 ***  |
| IUCN category  |  1.18672 |   0.33469  |   3.546 | 0.000391 *** |

### Coral depth map

@@

>  __Fig. S3.__ IUCN-defined range of _Oculina varicosa_, a typical photosynthetic coral, plotted against a 200 m bathymetry contour. According to IUCN, _O. varicosa_ can be found to depths of 152 m.  Maps of many other coral species and other reef-associated organisms followed similar patterns.

### AquaMaps threshold examination

@@

AquaMaps distribution maps indicate "probability of occurrence" within each 0.5° cell, with values ranging from zero to one, rather than a simple present/absent value as indicated by IUCN maps.  Many studies convert this AquaMaps probability to a simple presence value by assigning a threshold value.

To examine the effect of different presence threshold selections on the represented range of a species, we varied the threshold from 0% to 100% and calculated the average species range relative to a zero threshold.

@@

>  __Fig. S4.__ AquaMaps distribution map extent remaining after applying a presence threshold. A higher threshold constrains an analysis to cells with near certainty of occurrence, while a low threshold captures larger areas of increasingly marginal suitability. (A) A 40% threshold applied to all species in the AquaMaps dataset shows a mean loss of 28% of total range compared to a 0% threshold, with a wide distribution in which some species lose nearly all of their apparent range.  (B) Median remaining extent at increments of presence threshold. Dark green ribbon includes 25% to 75% quantiles of remaining species range, while lighter ribbon includes 5% to 95% quantiles.  Incrementing the presence threshold from 0.00 to 1.00 for the entire AquaMaps dataset, the shallow downward trend indicates a low but consistent sensitivity of apparent species range to threshold choice, with no surprising tradeoffs that could suggest an "optimal" threshold.

@@
